### 2022-01-17 manuscript/methodology.tex
correcao invariancia
revisao geophysics inclusas

Let $\mathbf{d}^{o}$ be the $N \times 1$ observed data vector, whose $i$th element 
is the total-field anomaly $d^{o}_{i}$ produced by arbitrarily magnetized sources
at the position $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $, 
of a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
pointing to north, east and down, respectively.
We consider that the total-field anomaly data $d^{o}_{i}$ represent the discrete
values of a harmonic function. Besides, we consider that the main geomagnetic field 
direction at the study area can be defined by the unit vector
\begin{equation}
\hat{\mathbf{F}} = \begin{bmatrix}
F_x \\
F_y \\
F_z
\end{bmatrix} =
\begin{bmatrix}
\cos(I_{0}) \, \cos(D_{0}) \\
\cos(I_{0}) \, \sin(D_{0}) \\
\sin(I_{0})
\end{bmatrix} \: ,
\label{eq:unit_vector_F}
\end{equation}
with constant inclination $I_{0}$ and declination $D_{0}$.
In this case, $d^{o}_{i}$ can be approximated by the predicted total-field anomaly
\begin{equation}
\Delta T_{i} = \sum_{j=1}^{M} \, p_{j} a_{ij} \: ,
\label{eq:integral-sum_mag}
\end{equation}
which describes the magnetic induction exerted, at the observation point $(x_{i}, y_{i}, z_{i})$,
by a discrete layer of $M$ dipoles (equivalent sources) defined on the horizontal plane $z = z_{c}$, 
where $p_{j}$ is the magnetic moment intensity (in A~m~$^{2}$)~of the $j$th dipole, 
that has unit volume and is located at the point $(x_{j}, y_{j}, z_{c})$. In equation
\ref{eq:integral-sum_mag}, $a_{ij}$ is the harmonic function
\begin{equation}
a_{ij}
= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{u}} \: ,
\label{eq:aij_mag}
\end{equation}
the unit vector
\begin{equation}
\hat{\mathbf{u}} = \begin{bmatrix}
u_x \\
u_y \\
u_z
\end{bmatrix} =
\begin{bmatrix}
\cos(I) \, \cos(D) \\
\cos(I) \, \sin(D) \\
\sin(I)
\end{bmatrix} \: ,
\label{eq:u_hat}
\end{equation}
defines the magnetization direction of all dipoles, with constant inclination $I$ and declination $D$,
$\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, $c_{m} = 10^{9}$ is a factor that transforms
the magnetic induction from Tesla (T) to nanotesla (nT) and $\mathbf{H}_{ij}$ is a $3 \times 3$ matrix 
\begin{equation}
\mathbf{H}_{ij} = \begin{bmatrix}
h^{xx}_{ij} & h^{xy}_{ij} & h^{xz}_{ij} \\
h^{xy}_{ij} & h^{yy}_{ij} & h^{yz}_{ij} \\
h^{xz}_{ij} & h^{yz}_{ij} & h^{zz}_{ij}
\end{bmatrix} \: ,
\label{eq:Hij}
\end{equation}
where 
\begin{equation}
h^{\alpha\beta}_{ij} = 
\begin{cases}
\frac{3 \left( \alpha_{i} - \alpha_{j} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: , \quad \alpha = \beta \\
\frac{3 \left( \alpha_{i} - \alpha_{j} \right) \left( \beta_{i} - \beta_{j} \right)}{r_{ij}^{5}} \: , \quad \alpha \ne \beta
\end{cases} \: , \quad \alpha, \beta = x, y, z \: ,
\label{eq:hij_alpha_beta}
\end{equation}
are the second derivatives of the inverse distance function
\begin{equation}
\frac{1}{r_{ij}} = 
\frac{1}{\sqrt{\left(x_{i} - x_{j} \right)^{2} + 
\left(y_{i} - y_{j} \right)^{2} + \left(z_{i} - z_{c} \right)^{2}}}
\label{eq:1_rij}
\end{equation}
with respect to the coordinates of the observation point $(x_{i}, y_{i}, z_{i})$.

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
as the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
\mathbf{A}^{\top} \mathbf{d}^{o} \: .
\label{eq:normal-equations}
\end{equation}
Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
factor $\mathbf{G}$ of matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
systems \citep[][ p. 262]{golub-vanloan2013}:
\begin{equation}
\begin{split}
\mathbf{G} \mathbf{w} &= \mathbf{A}^{\top}\mathbf{d}^{o} \\
\mathbf{G}^{\top} \tilde{\mathbf{p}} &= \mathbf{w}
\end{split} \: ,
\label{eq:classical-method}
\end{equation}
where $\mathbf{w}$ is a dummy variable.
This approach to estimate the parameter vector will be 
referenced throughout this work as the \textit{classical method}.
The computational cost associated with the classical method can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and 
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is formed by one dipole right below each observation point, at the constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
\label{eq:xi}
\end{equation}
and
\begin{equation}
y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the lower limits for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

By using equations \ref{eq:xi}--\ref{eq:l-y-oriented} to define the coordinates $x_{i}$ and 
$y_{i}$ of the observation points and $x_{j}$ and $y_{j}$ of the equivalent sources, we can
rewrite the elements $h^{\alpha\beta}_{ij}$ (equation \ref{eq:hij_alpha_beta}) of matrix 
$\mathbf{H}_{ij}$ (equation \ref{eq:Hij}) as follows:
\begin{equation}
h^{xx}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hxx_regular}
\end{equation}
\begin{equation}
h^{yy}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hyy_regular}
\end{equation}
\begin{equation}
h^{zz}_{ij} = 
\frac{3 \Delta_{z}^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hzz_regular}
\end{equation}
\begin{equation}
h^{xy}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)\left( \Delta l_{ij} \, \Delta_{y} \right)}{r_{ij}^{5}} \: ,
\label{eq:hxy_regular}
\end{equation}
\begin{equation}
h^{xz}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right) \Delta_{z}}{r_{ij}^{5}}
\label{eq:hxz_regular}
\end{equation}
and
\begin{equation}
h^{yz}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right) \Delta_{z}}{r_{ij}^{5}} \: ,
\label{eq:hyz_regular}
\end{equation}
where $\Delta_{z} = z_{c} - z_{0}$, 
\begin{equation}
\Delta k_{ij} = \frac{x_{i} - x_{j}}{\Delta_{x}} = k(i) - k(j) \: ,
\label{eq:Delta_kij}
\end{equation}
\begin{equation}
\Delta l_{ij} = \frac{y_{i} - y_{j}}{\Delta_{y}} = l(i) - l(j)
\label{eq:Delta_lij}
\end{equation}
and 
\begin{equation}
\frac{1}{r_{ij}} = 
\frac{1}{\sqrt{\left( \Delta k_{ij} \, \Delta_{x} \right)^{2} + \left( \Delta l_{ij} \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
\label{eq:1_rij_regular}
\end{equation}
Note that the integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ (equations 
\ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}) defining $\Delta k_{ij}$ (equation
\ref{eq:Delta_kij}), $\Delta l_{ij}$ (equation \ref{eq:Delta_lij}) and 
$\tfrac{1}{r_{ij}}$ (equation \ref{eq:1_rij_regular}) assume different 
forms depending on the grid orientation.
Despite of that, it can be shown that
\begin{equation}
\Delta k_{ij} = - \Delta k_{ji} \: ,
\label{eq:Delta_kij_symmetry}
\end{equation}
\begin{equation}
\Delta l_{ij} = - \Delta l_{ji}
\label{eq:Delta_lij_symmetry}
\end{equation}
and 
\begin{equation}
\frac{1}{r_{ij}} = \frac{1}{r_{ji}}
\label{eq:1_rij_symmetry}
\end{equation}
for any grid orientation.

@@

By using equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular} to compute 
$a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}), we can show that 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}) assume
well-defined structures that can be conveniently
represented with \textit{block indices} $q$ and $p$ \citep{takahashi2020convolutional}.
These indices are defined by the integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ 
(equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}), in terms of the indices $i$ 
of the observation points $(x_{i}, y_{i}, z_{0})$ and $j$ of the equivalent sources
$(x_{j}, y_{j}, z_{c})$.
For $x$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{y}$, $P = N_{x}$ 
and the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta l_{ij}
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta k_{ij} \: ,
\label{eq:p-x-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-x-oriented} and \ref{eq:l-x-oriented}.
For $y$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{x}$, $P = N_{y}$ and 
the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta k_{ij}
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta l_{ij} \: ,
\label{eq:p-y-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}.
Equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented} show that $q$ varies from $-Q+1$
to $Q-1$ and $p$ from $-P+1$ to $P-1$, regardless of the grid orientation. They differ 
from those presented by \citet{takahashi2020convolutional} due to the absence of the module.

@@

	8 - Repeat previous steps until convergence.

@@

Setting a convergence criteria based on the minimum tolerance of the residuals is a good 
option to carry out this algorithm efficiently and still obtaining very good results. 
Another possibility is to set an invariance to the Euclidean norm of residuals between 
iterations, which would increase algorithm runtime, but with smaller residuals. 
We chose the first option, as we achieve satisfactory results.

@@

In this sections we compare the efficiency of the classical (equation \ref{eq:classical-method}), 
standard CGLS (Algorithm \ref{al:std-cgls-algorithm}) and the convolutional equivalent 
layer method (Algorithm \ref{al:std-cgls-algorithm} with matrix-vector products computed 
according to Appendix A). To do this, we compute the total number of 
\textit{flops} associated to them \citep[][ p. 12]{golub-vanloan2013}.

@@

To compute the flops count of our method, we need only to replace the flops associated with 
matrix-vector products in the standard CGLS method by those associated with
2D convolution defined in Appendix A, which consists of $\kappa  \, 4 N \log_2(4N)$ flops to
compute the 2D FFT for each matrix $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ (equation 
\ref{eq:L_alpha_beta}); $\kappa  \, 4 N \log_2(4N)$ flops to compute 
$\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P}$ via 2D FFT; $24 \, N$ flops to compute the 
Hadamard product $\mathbf{L} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right)$; 
and $\kappa  \, 4 N \log_2(4N)$ flops to compute the IDFT in equation 
\ref{eq:2d-discrete-convolution-complete}. We use $\kappa = 5$ for the \emph{radix-2} algorithm
\citep[][ p.~15]{vanloan1992}. By replacing these flops into Algorithm \ref{al:std-cgls-algorithm},
we obtain the complete number of flops
\begin{equation}
f_{conv} =  \kappa  \, 16 N \log_2(4 N) + 24 N + it \, (\kappa  \, 16 N \log_2 (4 N) + 60 N) \: .
\label{eq:flops-convolutional-method}
\end{equation}

>>>>

Let $\mathbf{d}^{o}$ be the $N \times 1$ observed data vector $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $,
whose $i$th element is the total-field anomaly $d^{o}_{i}$ produced by arbitrarily magnetized sources
aranged in a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
pointing to north, east and down, respectively.
We consider that the total-field anomaly data $d^{o}_{i}$ represent the discrete
values of a harmonic function. Besides, we consider that the main geomagnetic field 
direction is constant at the limited study area and can be defined by the unit vector
\begin{equation}
\hat{\mathbf{F}} = \begin{bmatrix}
F_x \\
F_y \\
F_z
\end{bmatrix} =
\begin{bmatrix}
\cos(I_{0}) \, \cos(D_{0}) \\
\cos(I_{0}) \, \sin(D_{0}) \\
\sin(I_{0})
\end{bmatrix} \: ,
\label{eq:unit_vector_F}
\end{equation}
with constant inclination $I_{0}$ and declination $D_{0}$.
In this case, $d^{o}_{i}$ can be approximated by the predicted total-field anomaly
\begin{equation}
\Delta T_{i} = \sum_{j=1}^{M} \, p_{j} a_{ij} \: ,
\label{eq:integral-sum_mag}
\end{equation}
which describes the magnetic induction exerted at the observation point,
by a discrete layer of $M$ dipoles (equivalent sources) defined on the horizontal plane $z = z_{c}$, 
where $p_{j}$ is the magnetic moment intensity (in A~m~$^{2}$)~of the $j$th dipole, 
that has unit volume and is located at the point $(x_{j}, y_{j}, z_{c})$. In equation
\ref{eq:integral-sum_mag}, $a_{ij}$ is the harmonic function
\begin{equation}
a_{ij}
= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{u}} \: ,
\label{eq:aij_mag}
\end{equation}
the unit vector
\begin{equation}
\hat{\mathbf{u}} = \begin{bmatrix}
u_x \\
u_y \\
u_z
\end{bmatrix} =
\begin{bmatrix}
\cos(I) \, \cos(D) \\
\cos(I) \, \sin(D) \\
\sin(I)
\end{bmatrix} \: ,
\label{eq:u_hat}
\end{equation}
defines the magnetization direction of all dipoles, with constant inclination $I$ and declination $D$,
$\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, $c_{m} = 10^{9}$ is a factor that transforms
the magnetic induction from Tesla (T) to nanotesla (nT) and $\mathbf{H}_{ij}$ is a $3 \times 3$ matrix 
\begin{equation}
\mathbf{H}_{ij} = \begin{bmatrix}
h^{xx}_{ij} & h^{xy}_{ij} & h^{xz}_{ij} \\
h^{xy}_{ij} & h^{yy}_{ij} & h^{yz}_{ij} \\
h^{xz}_{ij} & h^{yz}_{ij} & h^{zz}_{ij}
\end{bmatrix} \: ,
\label{eq:Hij}
\end{equation}
where 
\begin{equation}
h^{\alpha\beta}_{ij} = 
\begin{cases}
\frac{3 \left( \alpha_{i} - \alpha_{j} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: , \quad \alpha = \beta \\
\frac{3 \left( \alpha_{i} - \alpha_{j} \right) \left( \beta_{i} - \beta_{j} \right)}{r_{ij}^{5}} \: , \quad \alpha \ne \beta
\end{cases} \: , \quad \alpha, \beta = x, y, z \: ,
\label{eq:hij_alpha_beta}
\end{equation}
are the second derivatives of the inverse distance function
\begin{equation}
\frac{1}{r_{ij}} = 
\frac{1}{\sqrt{\left(x_{i} - x_{j} \right)^{2} + 
\left(y_{i} - y_{j} \right)^{2} + \left(z_{i} - z_{c} \right)^{2}}}
\label{eq:1_rij}
\end{equation}
with respect to the coordinates of the observation point $(x_{i}, y_{i}, z_{i})$.

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
as the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
\mathbf{A}^{\top} \mathbf{d}^{o} \: .
\label{eq:normal-equations}
\end{equation}
Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
factor $\mathbf{G}$ of the positive-definite matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
systems \citep[][ p. 262]{golub-vanloan2013}:
\begin{equation}
\begin{split}
\mathbf{G} \mathbf{w} &= \mathbf{A}^{\top}\mathbf{d}^{o} \\
\mathbf{G}^{\top} \tilde{\mathbf{p}} &= \mathbf{w}
\end{split} \: ,
\label{eq:classical-method}
\end{equation}
where $\mathbf{w}$ is a temporarily variable.
This approach to estimating the parameter vector will be 
referenced throughout this work as the \textit{classical method}.
The computational cost associated with the classical method can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and 
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is aranged such that one dipole is located right below each observation point, at a constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
\label{eq:xi}
\end{equation}
and
\begin{equation}
y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the start grid coordinates for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

By using equations \ref{eq:xi}--\ref{eq:l-y-oriented} to define the coordinates $x_{i}$ and 
$y_{i}$ of the observation points and $x_{j}$ and $y_{j}$ of the equivalent sources, we can
rewrite the elements $h^{\alpha\beta}_{ij}$ (equation \ref{eq:hij_alpha_beta}) of matrix 
$\mathbf{H}_{ij}$ (equation \ref{eq:Hij}) as follows:
\begin{equation}
h^{xx}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hxx_regular}
\end{equation}
\begin{equation}
h^{yy}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hyy_regular}
\end{equation}
\begin{equation}
h^{zz}_{ij} = 
\frac{3 \Delta_{z}^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hzz_regular}
\end{equation}
\begin{equation}
h^{xy}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)\left( \Delta l_{ij} \, \Delta_{y} \right)}{r_{ij}^{5}} \: ,
\label{eq:hxy_regular}
\end{equation}
\begin{equation}
h^{xz}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right) \Delta_{z}}{r_{ij}^{5}}
\label{eq:hxz_regular}
\end{equation}
and
\begin{equation}
h^{yz}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right) \Delta_{z}}{r_{ij}^{5}} \: ,
\label{eq:hyz_regular}
\end{equation}
where $\Delta_{z} = z_{c} - z_{0}$, 
\begin{equation}
\Delta k_{ij} = \frac{x_{i} - x_{j}}{\Delta_{x}} = k(i) - k(j) \: ,
\label{eq:Delta_kij}
\end{equation}
\begin{equation}
\Delta l_{ij} = \frac{y_{i} - y_{j}}{\Delta_{y}} = l(i) - l(j)
\label{eq:Delta_lij}
\end{equation}
and 
\begin{equation}
\frac{1}{r_{ij}} = 
\frac{1}{\sqrt{\left( \Delta k_{ij} \, \Delta_{x} \right)^{2} + \left( \Delta l_{ij} \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
\label{eq:1_rij_regular}
\end{equation}
Note that the integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ (equations 
\ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}) defining $\Delta k_{ij}$ (equation
\ref{eq:Delta_kij}), $\Delta l_{ij}$ (equation \ref{eq:Delta_lij}) and 
$\tfrac{1}{r_{ij}}$ (equation \ref{eq:1_rij_regular}) assume different 
forms depending on the defined grid orientation.
Despite of that, it can be shown that
\begin{equation}
\Delta k_{ij} = - \Delta k_{ji} \: ,
\label{eq:Delta_kij_symmetry}
\end{equation}
\begin{equation}
\Delta l_{ij} = - \Delta l_{ji}
\label{eq:Delta_lij_symmetry}
\end{equation}
and 
\begin{equation}
\frac{1}{r_{ij}} = \frac{1}{r_{ji}}
\label{eq:1_rij_symmetry}
\end{equation}
for any grid orientation.

@@

By using equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular} to compute 
$a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}), we can show that 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}) assume
well-defined structures that can be conveniently
represented with \textit{block indices} $q$ and $p$ \citep{takahashi2020convolutional}.
These indices are defined by the integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ 
(equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) in terms of the indices $i$ 
of the observation points $(x_{i}, y_{i}, z_{0})$ and $j$ of the equivalent sources
$(x_{j}, y_{j}, z_{c})$.
For $x$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{y}$, $P = N_{x}$ 
and the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta l_{ij}
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta k_{ij} \: ,
\label{eq:p-x-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-x-oriented} and \ref{eq:l-x-oriented}.
For $y$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{x}$, $P = N_{y}$ and 
the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta k_{ij}
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta l_{ij} \: ,
\label{eq:p-y-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}.
Equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented} show that $q$ varies from $-Q+1$
to $Q-1$ and $p$ from $-P+1$ to $P-1$, regardless of the grid orientation. Notice these equations differ 
from those presented by \citet{takahashi2020convolutional} by not being represented as the absolute values.

@@

	8 - Repeat previous steps until convergence ($\delta$).

@@

Setting a convergence criterion $\delta$ (Algorithm \ref{al:std-cgls-algorithm}) based on the minimum tolerance 
of the residuals is a good option to carry out this algorithm efficiently and still obtaining very good results. 
Another possibility is to set an invariance to the Euclidean norm of residuals between 
iterations, which would increase algorithm runtime, but with smaller residuals. 
We chose the latter option, as we could achieve better results.

@@

In this section we compare the efficiency of the classical (equation \ref{eq:classical-method}), 
standard CGLS (Algorithm \ref{al:std-cgls-algorithm}) and the convolutional equivalent 
layer method (Algorithm \ref{al:std-cgls-algorithm} with matrix-vector products computed 
according to Appendix A). To do this, we compute the total number of 
\textit{flops} associated to them \citep[][ p. 12]{golub-vanloan2013}.

@@

To compute the flops count of our method, we need only to replace the flops associated with 
matrix-vector products in the standard CGLS method by those associated with
2D convolution defined in Appendix A, which consists of $\kappa  \, 4 N \log_2(4N)$ flops to
compute the 2D FFT for each matrix $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ (equation 
\ref{eq:L_alpha_beta}); $\kappa  \, 4 N \log_2(4N)$ flops to compute 
$\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P}$ via 2D FFT; $24 \, N$ flops to compute the 
Hadamard product $\mathbf{L} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right)$; 
and $\kappa  \, 4 N \log_2(4N)$ flops to compute the IDFT (inverse discrete Fourier transform) in equation 
\ref{eq:2d-discrete-convolution-complete}. We use $\kappa = 5$ for the \emph{radix-2} algorithm
\citep[][ p.~15]{vanloan1992}. By replacing these flops into Algorithm \ref{al:std-cgls-algorithm},
we obtain the complete number of flops
\begin{equation}
f_{conv} =  \kappa  \, 16 N \log_2(4 N) + 24 N + it \, (\kappa  \, 16 N \log_2 (4 N) + 60 N) \: .
\label{eq:flops-convolutional-method}
\end{equation}
